Enhancing Adversarial Robustness for Deep Metric Learning

### 2021-10-22 robdml.tex
tex

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and the anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified hardness level.
%
HM is implemented as an objective to minimize the L-$2$ norm of the thresholded
difference between the hardness of the given sample triplet and the specified
hardness.
%
Mathematically, when the objective is optimized using Projected Gradient
Descent~\cite{madry}, the gradient of HM objective with respect to the
adversarial perturbation is the same as that of directly \emph{maximizing} the
triplet loss.
%
Thus, the optimization process of HM objective can be interpreted as a direct
\emph{maximization} process of the loss (indicating a direct strong adversary)
which stops halfway at the specified hardness level.
%
But, how hard should such ``specified level'' be?

We conjecture there is a certain intermediate level of hardness.

notable penalty in the model performance on benign (\ie, unperturbed) examples.

%
\oo{rev}
%
% inspiration.
Recall that a DML model is already prone to collapse with very hard but not
adversarially purturbed sample triplets~\cite{facenet}.
%
To overcome this issue, various triplet sampling strategies, such as
semi-hard~\cite{facenet} and distance~\cite{distance} are proposed to provide
effective negative examples to facilitate metric learning while avoiding model
collapse.
%
In the typical min-max adversarial training~\cite{madry}, the adversarial
example triplet created by maximizing the triplet loss value can incur model
collapse more easily than unperturbed sample triplets.
%
But what if we do not maximize the triplet loss value to the theoretical upper
bound, but merely increase it towards a value that could be caused by
a known, unperturbed and harder sample triplet?

>>>>

In this paper, we first define the ``\emph{hardness}'' of a sample triplet as
the difference between the anchor-positive distance and the anchor-negative
distance.
%
Then, Hardness Manipulation (HM) is proposed to adversarially perturb a given
sample triplet and increase its hardness into a specified destination hardness level.
%
HM is implemented as an objective to minimize the L-$2$ norm of the thresholded
difference between the hardness of the given sample triplet and the specified
destination hardness.
%
Mathematically, when the objective is optimized using Projected Gradient
Descent~\cite{madry}, the gradient of HM objective with respect to the
adversarial perturbation is the same as that of directly \emph{maximizing} the
triplet loss.
%
Thus, the optimization process of HM objective can be interpreted as a direct
\emph{maximization} process of the loss (indicating a direct strong adversary)
which stops halfway at the specified destination hardness level.
%
But, how hard should such ``destination hardness'' be?

% how hard? benign destination reference

Recall that the model is already prone to collapse with excessively hard
benign triplets, let alone adversarial examples.
%
Thus, intuitively, the destination hardness can be that of another benign
triplet which is harder than the given triplet (\eg, a Semihard~\cite{facenet}
triplet).
